---
title: "Cinci-Dayton RUG Meetup"
author: "Eugene Pyatigorsky"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook: 
    highlight: tango
    theme: cosmo
---
------

This notebook is available on:
https://github.com/epspi/02.28.2017_Cin-Day_RUG_sparklyr

The data is available at census.gov:
https://www.census.gov/econ/cfs/pums.html

A fully instructive tutorial is at:
http://spark.rstudio.com/


## Installation & Connection
`sparklyr` and `dplyr` is what we'll be using. Install Spark directly from R with the handy `spark_install()` function within `sparklyr`
```{r}
library(dplyr)
library(data.table)
library(sparklyr)
# spark_install(version = "1.6.2")
config <- spark_config()
config$sparklyr.gateway.port <-  5454
sc <- spark_connect(master = "local", config = config)
sc
src_tbls(sc)
```

## Importing Data
We need to feed data into the Spark cluster, whether by copying it from R objects or by using one of the filereader functions.  

#### copy_to
```{r}
system.time({
    ship <- fread("data/cfs_2012_pumf_csv.txt")
    ship_sp <- copy_to(sc, ship, "ship", overwrite = T)
})
src_tbls(sc)

```
  
  
#### spark_read_csv
```{r}
system.time(
    ship2_sp <- spark_read_csv(sc, "ship2", 
                           "data/cfs_2012_pumf_csv.txt")
)
```

We can't use `names` to figure out what's in the spark tables,
but `colnames` and `tbl_vars` work
```{r}
colnames(ship_sp)
tbl_vars(ship2_sp)
```


## Manipulating Data with dplyr

You can use **some** (all?) of the dplyr verbs and also SQL commands directly on spark tables. That's basically the whole point.

```{r}
ship_sp %>% count(ORIG_STATE)
```

Spark will defer calculation until you deliberately or implictely try to `collect` the data back as an R object. 

```{r}
head(ship_sp)
spark_log(sc, n = 3)
```

For example, we select some columns of interest and assign the resulting df to a new variables. However, as we see by calling `names` or `str` on the resulting variable, it's still a spark object, not a **collected** df.
```{r}
ship_values <- ship_sp %>% 
    select(contains("SHIPMT"))

names(ship_values)
```

## Machine Learning

The `sparklyr` API includes interfaces to way more `Spark ML` facilities than the previous API from Spark directly.

### Regression Example

Let's try running a simple linear regression of shipment values of some other numeric variables

```{r}
head(ship_values)
```

```{r}
lm_sp <- ship_values %>% 
    select(-SHIPMT_ID) %>% 
    ml_linear_regression(SHIPMT_VALUE ~ .)

dat <- ship %>% 
    select(contains("SHIPMT")) %>% 
    select(-SHIPMT_ID)

lm_R <- lm(SHIPMT_VALUE ~ ., data = dat)
    
```
```{r}
class(lm_sp)
class(lm_R)
```


```{r}
summary(lm_sp)
summary(lm_R)
```

